---
title: 'Reproducible Research Assignment #1'
author: "Seth Rivers"
date: "July 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## set working directory and load data
```{r}
setwd("C:/Users/srivers/Desktop/R_Workspace/Reproduce_Research/RepData_PeerAssessment1")
data<-read.csv("activity/activity.csv")
```


##turn in to Data Table
```{r}
library(data.table)
data<-data.table::data.table(data)
```

## Create Data set which excludes NAs

```{r}
data_no_nas<-data[complete.cases(data),]
```


##Aggregate steps by day

```{r}
steps_by_day<-data_no_nas[,list(Daily_steps=sum(steps)),by="date"]
```

##Create Histogram of Steps by day
```{r}
hist(steps_by_day$Daily_steps)
```


#Find Mean and Median
```{r}
Daily_steps_mean<-mean(steps_by_day$Daily_steps)
Daily_steps_median<-median(steps_by_day$Daily_steps)

Daily_steps_mean
Daily_steps_median

```

##Aggregate Steps by 5 minute interval
```{r}
steps_by_interval5<-data_no_nas[,list(interval_steps_avg=mean(steps)),by="interval"]
```

##Time Series Plot
```{r}
with(steps_by_interval5,plot(interval,interval_steps_avg,type = "l"))
```

##Find Interval with maximum number of steps
```{r}
max_steps<-max(steps_by_interval5$interval_steps_avg)
interval_w_max_steps<-steps_by_interval5[interval_steps_avg==max_steps,interval]
interval_w_max_steps
```




##Calculate total number of Missing Values
```{r}
missing_values_data<-data[!complete.cases(data)]
missing_values_count<-nrow(missing_values_data)
missing_values_count
```

##Impute Missing Values with the average with the overall interval average
```{r}
data_imputed<-merge( data, steps_by_interval5[,.(interval,interval_steps_avg)] ,all.x = TRUE)
data_imputed$steps<-ifelse(is.na(data_imputed$steps),data_imputed$interval_steps_avg,data_imputed$steps)
```



##Aggregate steps by day with imputed data

```{r}
steps_by_day_imputed<-data_imputed[,list(Daily_steps=sum(steps)),by="date"]
```

##Create Histogram of Steps by day of imputed data
```{r}
hist(steps_by_day_imputed$Daily_steps)
```


##Find Mean and average of imputed data

```{r}
Daily_steps_mean_imputed<-mean(steps_by_day_imputed$Daily_steps)
Daily_steps_median_imputed<-median(steps_by_day_imputed$Daily_steps)

Daily_steps_mean_imputed>Daily_steps_mean
Daily_steps_mean_imputed==Daily_steps_mean
Daily_steps_median_imputed>Daily_steps_median


```

There is no difference between the average steps by data between the original data and the imputed data
However the median of the imputed data is slightly larger

##Create Weekday Weekend Factor
```{r}
data_imputed$day_of_week<-weekdays(as.Date(data_imputed$date),TRUE)
data_imputed$day_type<-factor(ifelse(data_imputed$day_of_week %in% c('Sat','Sun'),'weekend','weekday'))

```


##Aggregate data by interval/day_type for panel plot
```{r}
steps_by_interval5_imputed<-data_imputed[,list(interval_steps_avg=mean(steps)),by=c("interval","day_type")]
```


##Create Panel Plot of average number of steps by interval for weekdays and weekends
```{r}
library(ggplot2)
ggplot(steps_by_interval5_imputed,aes(interval,interval_steps_avg))+facet_wrap(~day_type)+geom_line()


```




















